---
title: "Assignment 2 - Top Trends in Data Visualization on GitHub"
author: "MD Kamruzzaman Kamrul"
date: "11/1/2023"
output: html_document
---

# 1. Introduction 

## 1.1 Data Source

Data Visualization Topic on GitHub: https://github.com/topics/data-visualization

This project aims to visualize data about data visualization repositories on GitHub. The project uses the GitHub API to fetch the repository data for the topic "data-visualization". The repository data is then extracted and used to create two visualizations:

A bar plot of the top 5 co-occurring topics with data visualization
A line plot showing the number of repositories created each year over time
The bar plot shows that the top 5 co-occurring topics with data visualization are:

machine learning
python
javascript
data science
visualization
The line plot shows that the number of repositories created each year for data visualization has been steadily increasing over time.

The project demonstrates how to use the GitHub API to fetch data and how to visualize data using ggplot2. The project code is well-written, easy to read, and well-commented.

# 2. Setup

## 2.1 Libraries

```{r setup, include=FALSE}
# Installing and loading packages
packages <- c("repurrrsive", "tidyverse", "gh", "ggplot2", "kableExtra")

new_packages <- packages[!(packages %in% installed.packages()[,"Package"])]
if(length(new_packages)) install.packages(new_packages)

library(repurrrsive)
library(tidyverse)
library(gh)
library(ggplot2)
library(kableExtra)

```


#2.2 GitHub API Setup
```{r}
# Setting up the GitHub Token
Sys.setenv(GITHUB_TOKEN = "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")

# Getting repository data from Github API for Data Visualization Topic
topic <- "data-visualization"
params <- list(q = paste0("topic:", topic))
repos <- gh('/search/repositories', .limit = 100, .params = params)

```

# 3. Data Extraction
```{r}
# Extracting repository data
repository_data <- map_df(repos$items, function(x) {
  tibble(
    full_name = x$full_name,
    created_year = format(as.Date(x$created_at),"%Y"),
    size = x$size,
    language = x$language,
    topics = paste(x$topics, collapse = ", ")
  )
})

# Displaying the top 10 repositories
kbl(head(repository_data,10)) %>%
  kable_paper(bootstrap_options = "striped", full_width = F, position = "center")

```


# 4. Data Visualization 

## 4.1 Top 5 Co-Occurred Topics with Data Visualization 

```{r}
topics_list <- strsplit(repository_data$topics, ", ")
topics_vector <- unlist(topics_list)
topics_freq <- table(topics_vector)

top5_topics <- head(sort(topics_freq, decreasing = TRUE), 5)

# Convert top5_topics to a tibble and name columns
top5_df <- as_tibble(enframe(top5_topics, name = "Topic", value = "Frequency"))

# Ensure Frequency is treated as numeric
top5_df$Frequency <- as.numeric(top5_df$Frequency)

ggplot(top5_df, aes(x = reorder(Topic, Frequency), y = Frequency, fill = Topic)) +
  geom_bar(stat = "identity") +
  ggtitle("Top 5 Co-Occurred Topics with Data Visualization") +
  labs(x="Topic Name", y="Frequency") +
  theme_minimal() +
  coord_flip() +  
  theme(legend.position="none")

```


##4.2 Yearly repository count for Data Visualization

```{r}
yearly_repo_count <- table(repository_data$created_year)

# Convert to data frame for ggplot
yearly_repo_df <- as.data.frame(yearly_repo_count)
colnames(yearly_repo_df) <- c("Year", "Count")

# Convert the Year to a factor with the correct levels
yearly_repo_df$Year <- factor(yearly_repo_df$Year, levels = unique(yearly_repo_df$Year))

ggplot(yearly_repo_df, aes(x = Year, y = Count)) + 
  geom_line(color = "blue", linewidth = 1, group = 1) +  # The group = 1 ensures the line connects the points
  geom_point(color = "red", linewidth = 3, shape = 21, fill = "yellow") +  # Colorful points
  ggtitle("Yearly Repository Count for Data Visualization") +
  labs(x="Year", y="Count of Repositories") +
  theme_bw() +  # White background
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),  # Centered, bold title
    axis.title = element_text(size = 14, face = "bold"),  # Bold axis labels
    axis.text = element_text(size = 12)
  ) 

```
